### 2014-02-21 paper/abstract.tex
moar draft
  * figure of tube design space.  text about it.  some text about
    the tool, some intro.  abstract.  placeholders for most figures.
    basically just more text.

\begin{abstract}
Using different types of tubes with different topologies and different meda inside, we can make cool printable interfaces wtih a variety of input and output mechanisms.
\end{abstract}

>>>>

\begin{abstract}
3D printers offer extraordinary flexibilty in prototyping the shape and mechanical function of objects.  However, in spite of recent work, a 3D printer's role in prototyping \emph{interactivity} is still not clear.  While macro-scale digital fabrication devices cannot manufacture electronics in-place, there have been many clever approaches to redirecting interaction onto 3D prints.  We describe the design space of tubes and hollow chambers for interaction design: there are a variety of types, topologies, and inserted media for tubes that can be leveraged to create diverse inputs and outputs.  We present a technique and design tool for routing tubes of various topologies through the interior of 3D printed parts.  Our design tool is integrated into a 3D model manipulation program.  It allows users to select begin and end points for tubes, then uses A* path routing and physics-based simulation to minimize the bending energy of routed paths.  We present several totally tubular prototypes we created using this tool to show its flexibility and potential, as well as to explore new points in the tube design space.
\end{abstract}

### 2014-02-21 paper/introduction.tex
moar draft
  * figure of tube design space.  text about it.  some text about
    the tool, some intro.  abstract.  placeholders for most figures.
    basically just more text.

\section{Introduction}
3D printers on the rise, different kinds of materials, don't want to make knick knacks, personalizable interfaces, cheap components, etc.

>>>>

\begin{figure}[h]
\centering
    \includegraphics[width=3.4in]{figures/series-of-tubes.jpg}
\caption{Here's a teaser figure of some cool stuff we did. Probably it should show one of our example objects, or we can make it really big (figure*) and show them all!}
\label{fig:teaser}
\end{figure}

\section{Introduction}
Makers, as well as professional designers, leverage 3D printers as tools for design work.  A wide array of objects, ranging from bicycle helmets to jewelry to video game controllers, are now prototyped or even manufactured using these machines.  However, most devices fabricated by 3D printers are passive: accessible printers are not yet capable of creating integrated active systems.

Willis, et al., ``envision a future world where interactive devices can be printed rather than assembled; a world where a device with active components is created as a single object, rather than a case enclosing circuit boards and individually assembled parts" \cite{Willis-printedoptics}.  This is a vision we eagerly share: 3D printers are capable of greating arbitrary geometries not feasible to manufacture using traditional processes, and we see these capabilities being underutilized by makers and designers.  We see many opportunitities to increase the interactivity of 3D prints using today's printer technology.

Using specially modified printers or extra machinery, it is possible to create electronics on the surface of 3D prints \cite{optomec} \cite{Sells-reprap}.  These techniques, however, require a high capital investment and technical expertise.  In addition, they lack flexibility: the printed circuits must be routed in 2D to be created on the objects' manifold exteriors, and they can only be used to create electronic circuits. \valkyrie{what?}

In addition to electronics, many other means of interactivity exist.

\valkyrie{Points to hit: there is more than electronics to worry about, like fluids and the organic haptic feedback they afford; tangible interfaces are important (do I have to say this every time??); assembly sucks}

### 2014-02-21 paper/relatedwork.tex
moar draft
  * figure of tube design space.  text about it.  some text about
    the tool, some intro.  abstract.  placeholders for most figures.
    basically just more text.

\begin{itemize}
\item \cite{Sarik-tracebrush} - this uses an airspray machine to add conductive paths to surface of 3D prints
\item \cite{Harrison-buttons} - fabricate latex + acrylic buttons and pressurize with air
\item \cite{Hashimoto-squirming} - hold a speaker in your hands, and air pressure changes make it feel like you're holidng a living, squirming thing
\item \cite{Kim-inflatablemouse} - a basic mouse, but it inflates so you can store it and also use it more reasonably than a flat mouse
\item \cite{Ono-touchandactivate} - the Touch\'{e} with sound paper from last year's UIST
\item \cite{Iwata-volflex} - a display made up of many balloons that inflate and deflate to change the shape
\item \cite{Sells-reprap} - the RepRap people using a syringe of hot solder to squirt flat circuits into flat channels
\item \cite{Majidi-curvature} - inject liquid metal into really thin tubes in a soft substrate, sense stretching by changing resistance
\item \cite{Navarrette-gps} - I'm a little unclear on what they did, but they fabbed something with a "3D circuitboard" that has a bypass that goes into 3D.  they don't offer a routing algorithm or anything, though.
\item \cite{Optomec} - fancy machines that spray conductive film
\item \cite{Slyper-pressure} - robots that are made of squishy stuff where air pressure changes are sensed.  input components designed to react to different manipulations (pushing, squeezing, twisting, etc.)
\item \cite{icecats-conductivepaint} - an instructable I followed to make my first batch of conductive paint
\item \cite{Follmer-jamming} - Jamming User Interfaces
\item \cite{Park-microchannels} - similar to injection of liquid metal, above
\item \cite{Slyper-shape} - creating silicone bendy things with embedded electronics to sense flexing, stretching, etc., supported by those shapes.
\item \cite{Yao-pneui} - PneUIs, creating interfaces with pneumatics
\item \cite{Willis-infrastructs} - InfraStructs, identification of 3d printed objects using Terahertz imaging
\item \cite{Willis-printedoptics} - Printed Optics.  doing cool stuff with clear material (touch sensing and display)
\item \cite{Sodhi-aireal} - air vortex generation in free space.  air haptics.
\item \cite{Savage-midas} - Midas.  routing in 2D to connect up capacitive sensors
\item \cite{Savage-sauron} - Sauron.  computer vision of mechanical components obviates electronics installation.
\item \cite{Sato-touche} - Touch\'{e}.  It's like Touch\'{e} with sound, but without sound.  SFCS.
\end{itemize}

@@

\subsection{Printing Electronics}

\subsection{Fluids For Interaction}

\subsection{Routing}

\subsection{The Design Space}

\begin{table*}[t]
\caption{Existing systems are written in regular font.  What's in \emph{italic} are things that I think we could make.  I don't know what else might go in the blank spaces... looking for suggestions!  Also, I'm not confident this breakdown is the most clear: for example, the "liquids" category includes the copper paint, which functions as wires ("threadables"), but takes the form of a liquid...}
\begin{tabular}{ r || p{3cm} | p{3cm} | p{3cm} | p{3cm} }
  \hline
   & Gas & Liquid & Particles & Threadables\\ \hline \hline
  Visual & \parbox{3cm}{PneUIs, \\Harrison latex thingy, \\ \emph{smoke display}} & \parbox{3cm}{Splash Controllers, \\ \emph{paint mixer}} &\parbox{3cm}{\emph{clear thing with popcorn}} &\parbox{3cm}{\emph{neon sign}} \\ \hline
  Aural & \parbox{3cm}{\emph{resonance}} & \parbox{3cm}{\emph{bubbling/splashing noises}} &\parbox{3cm}{\emph{CNC maracas}} &\parbox{3cm}{} \\ \hline
  Tactile/Haptic & \parbox{3cm}{PneUIs,\\ \emph{haptic textures}} & \parbox{3cm}{Splash Controllers,\\ \emph{warm/cold liquid}} &\parbox{3cm}{Jamming UIs, \\ \emph{sparse particle haptic textures}} &\parbox{3cm}{Otherlab robots} \\ \hline
  Olfactory/Gustatory & \parbox{3cm}{\emph{scents}} & \parbox{3cm}{\emph{smoothie mixer!}} &\parbox{3cm}{} &\parbox{3cm}{} \\ \hline
  Touch Input & \parbox{3cm}{Harrison latex thingy} & \parbox{3cm}{Harrison SFCS, \\ \emph{injectable capacitive sensors}} &\parbox{3cm}{Jamming UIs} &\parbox{3cm}{\emph{capacitance on wires}} \\ \hline
  Pressure Input & \parbox{3cm}{Slyper printed doodads} & \parbox{3cm}{\emph{capacitance, flow meter}} &\parbox{3cm}{} &\parbox{3cm}{} \\ \hline
  Other Input & \parbox{3cm}{Slyper (twist, bend, flex, etc.), \\ \emph{printed flex sensors}} & \parbox{3cm}{\emph{use traditional components}} &\parbox{3cm}{} &\parbox{3cm}{\emph{use traditional components}} \\ \hline
\end{tabular}
\end{table*}
\

>>>>

Our work is related to three existing research areas: fabrication, interaction techniques, and routing.

@@

Previous work has investigated the integration of interactivity into the fabrication process.  \cite{Willis-printedoptics} used light pipes, solid clear tubes, to create integrated displays.  \cite{Savage-sauron} investigated the use of computer vision to track interactions with physical mechanisms.
\begin{itemize}
\item \cite{Savage-sauron} - Sauron.  computer vision of mechanical components obviates electronics installation.
\item \cite{Willis-infrastructs} - InfraStructs, identification of 3d printed objects using Terahertz imaging
\item \cite{Willis-printedoptics} - Printed Optics.  doing cool stuff with clear material (touch sensing and display)
\item \cite{Slyper-pressure} - robots that are made of squishy stuff where air pressure changes are sensed.  input components designed to react to different manipulations (pushing, squeezing, twisting, etc.)
\item \cite{Navarrette-gps} - I'm a little unclear on what they did, but they fabbed something with a "3D circuitboard" that has a bypass that goes into 3D.  they don't offer a routing algorithm or anything, though.
\end{itemize}

\subsubsection{Printing Electronics}
\begin{itemize}
\item \cite{Sarik-tracebrush} - this uses an airspray machine to add conductive paths to surface of 3D prints
\item \cite{Optomec} - fancy machines that spray conductive film
\item \cite{Sells-reprap} - the RepRap people using a syringe of hot solder to squirt flat circuits into flat channels
\item \cite{icecats-conductivepaint} - an instructable I followed to make my first batch of conductive paint
\end{itemize}

\subsection{Routing}
\begin{itemize}
\item \cite{Savage-midas} - Midas.  routing in 2D to connect up capacitive sensors
\item \cite{Park-microchannels} - similar to injection of liquid metal, above
\item \cite{Majidi-curvature} - inject liquid metal into really thin tubes in a soft substrate, sense stretching by changing resistance
\end{itemize}

\subsection{Interaction Techniques}
\begin{itemize}
\item \cite{Sato-touche} - Touch\'{e}.  It's like Touch\'{e} with sound, but without sound.  SFCS.
\item \cite{Harrison-buttons} - fabricate latex + acrylic buttons and pressurize with air
\item \cite{Sodhi-aireal} - air vortex generation in free space.  air haptics.
\item \cite{Yao-pneui} - PneUIs, creating interfaces with pneumatics
\item \cite{Slyper-shape} - creating silicone bendy things with embedded electronics to sense flexing, stretching, etc., supported by those shapes.
\item \cite{Iwata-volflex} - a display made up of many balloons that inflate and deflate to change the shape
\item \cite{Ono-touchandactivate} - the Touch\'{e} with sound paper from last year's UIST
\item \cite{Follmer-jamming} - Jamming User Interfaces
\item \cite{Kim-inflatablemouse} - a basic mouse, but it inflates so you can store it and also use it more reasonably than a flat mouse
\item \cite{Hashimoto-squirming} - hold a speaker in your hands, and air pressure changes make it feel like you're holidng a living, squirming thing
\end{itemize}

### 2014-02-21 paper/tubes.tex
moar draft
  * figure of tube design space.  text about it.  some text about
    the tool, some intro.  abstract.  placeholders for most figures.
    basically just more text.

\subsection{Types of Tubes}

\begin{itemize}
\item open - system to user, both ends open
\item return - system to system, both ends open
\item semi-closed - system to user, system side open, user side covered (e.g. elastic material)
\item fully closed - no openings on system or user side (e.g. air bubbles, resonance chambers)
\end{itemize}

\subsection{Features of Tubes}

\begin{itemize}
\item emphasize exterior connection points
\item emphasize interior design/path of tubes (for display)
\end{itemize}

\subsection{Media in Tubes}

\begin{itemize}
\item Gas (compressible).  Air with smells.  Air with fog.
\item Liquid (incompressible).  Clear, colored, conductive, heated, flavored (smoothie!).  Fill or coat.
\item Particles.  Sparse for haptic feedback.  Dense for jamming.
\item Threadables.  EL wire, conductive wire, fiberoptic cable, muscle wire.
\end{itemize}

\subsection{Design, Fabrication, Construction}

\subsubsection{Two Tools for Design}

\textbf{designing interior paths}

\textbf{designing exterior connection points}

\subsection{Inputs}

\subsubsection{Flexing}

Much like \cite{Slyper-shape}, we can sense flexing and bending of prints made on the Objet.  We can make prints on the Objet and tunnel through them, though, without making crazytastic silicone molds.

\subsubsection{Touch}
Capacitance (digital).

\subsubsection{Pressure}
Pressure (capacitance (ish) or fluid pressure).

\subsubsection{Tapping}
 Tapping (audio/resonance?) carried through particular tubes, like we talked about hard tubes in a soft thing.

\subsubsection{Other stuff?}
Could be.

\subsubsection{Traditional Components}
Obviously you can hook up traditional buttons, etc., the same way as always.  With copper paint instead of traditional wiring, we can share grounds, and we don't have to solder.

\subsection{Outputs}

\subsubsection{Visual}
EL wire.  Colored liquids.  Mechanical motion by pushing light stuff with air.

\subsubsection{Aural}
Resonance chambers.  Air cavity design for sound/amplification (see passive iPhone speakers).  I mean, this is basically just 3D printing instruments, which we know has been done.

\subsubsection{Haptic}
Compressible and incompressible fluids for actuation.  Recreation of PneUIs.  Tactile output.  Adding particles to add extra feedback.

\subsubsection{Olfactory/Gustatory}
Different chambers filled with different scented/flavored fluids.  We can mix them using pipes and pressure.

\subsection{Identity}
Resonance chambers for identification.

>>>>

\subsection{Tool for Tube Design}

We created a tool which novice designers can use to create tube-powered interfaces in arbitrary 3D models.  This tool allows users to brush over the surface of their model to either select exterior connection points of their tubes (see Figure \ref{fig:tool-process-interior}) or to author the tubes' interior paths (see Figure \ref{fig:tool-process-exterior}).  Once the user's selections are made, our tool creates an initial shortest-path routing using A* to estimate the routed distance between points.  This routing is used to create a rod; we run physics-based simulation steps on the rod to minimize its bending energy (and thereby minimize the bend radius of the tubes).  The resultant routing is subtracted from the user's mesh.  The modified mesh can then be 3D printed.

\subsubsection{Selection}

\textbf{Exterior Connection Points}

\begin{figure}[h!]
\centering
    \includegraphics[width=3.4in]{figures/series-of-tubes.jpg}
\caption{This figure has several sub-figures.  a) shows a model with exterior connection points brushed.  b) shows initial routing with A*.  c) shows our physics-based, energy-minimizing rod/tube, d) shows the printed object with the tube (with something in it, copper paint presumably)}
\label{fig:tool-process-exterior}
\end{figure}

Focus on shape of points and location of points.

\textbf{Designing Interior Paths}

\begin{figure}[h!]
\centering
    \includegraphics[width=3.4in]{figures/series-of-tubes.jpg}
\caption{This figure has several sub-figures.  a) shows a model with path shape and exterior connection point(s) brushed.  b) highlights the points which can't be tubed as drawn (if they are not connected, too tight, etc.).  c) shows our physics-based, energy-minimizing rod/tube and any path-smoothing that we do, d) shows the printed object with the tube (with something in it, EL wire presumably)}
\label{fig:tool-process-interior}
\end{figure}

take into account bend radius of desired material - no need, we just do the minimum-bending path

\subsubsection{Routing}

Our basic first-pass routing algorithm uses the A* routing algorithm \cite{Hart-Astar}.  The path cost in our implementation is based only on shortest distance between the starting and ending points, without weighting for distance from the surface.

\subsubsection{Physical Simulation}

\subsubsection{Mesh Modification}

\subsection{Fabrication Techniques}

\subsubsection{printing} - different strategies with Objet (all print-in-place) and Makerbot (may need to add things like balloons afterwards).  Ryan just got flexible material, we should see how stretchy it is...!  We could also consider assembleable things that are easier to create using parts that clip together... probably out of scope.

\subsubsection{hand tools} - post-fabrication modification is possible using hand tools.  We can mark the surface to show where conduits are and how deep.  I can also use this to test things beforehand.
